---
title:"EDA"
author:"Andrea Catucci"
date:"2025-05-27"
output:html_document
---
```{r setup, include = FALSE}
library(dplyr)
library(readxl)
library(sf)
library(terra)
library(spData)
library(spDataLarge)
library(tmap)
library(lubridate)
library(tidyr)
library(ggplot2)
library(zoo)
library(car)
library(purrr)
library(broom)
library(gstat)
library(sp)
library(spacetime)
library(furrr)
library(future)
library(future.apply)
library(arrow)
library(corrplot)
library(elevatr)
library(slider)
library(lightgbm)
library(Matrix)
library(suncalc)
```


# EDA

### First, add two new variables:
-Elevation.
-Rain lag in previous two hours.

#### Elevation
```{r altitude}
data <- readRDS('/home/andreacatucci/output_final_clean.RDS')

data_nogeom <- data %>%
mutate(
Longitude = st_coordinates(geometry)[, 1],
Latitude  = st_coordinates(geometry)[, 2],
Year      = year(datetime_hour),
Month     = month(datetime_hour),
Day       = day(datetime_hour),
Hour      = hour(datetime_hour),
Quarter   = quarter(datetime_hour)
) %>%
select(-geometry)

gsp_points <- data %>%
select(gsp_id, geometry) %>%
distinct() %>%
st_as_sf()

gsp_points_elev <- get_elev_point(gsp_points, src = "aws", z = 10)

gsp_elevations <- gsp_points_elev %>%
st_drop_geometry() %>%
select(gsp_id, elevation) %>%
rename(Elevation = elevation)

data_nogeom <- left_join(data_nogeom, gsp_elevations, by = "gsp_id")

rm(list = c("gsp_elevations", "gsp_points", "gsp_points_elev", "data"))
```


### HAS_RAINED variable to account for wet soil and panel conditions.
#### We add a binary variable that is 1 if it has rained in the previous 2 hours and the current (PRCP_AMT > 0), and 0 otherwise.

```{r rain_var}
data_nogeom <- data_nogeom %>%
arrange(gsp_id, datetime_hour) %>%
group_by(gsp_id) %>%
mutate(
HAS_RAINED = slide_lgl(
PRCP_AMT > 0,
.before = 2,
.after = 0,
.complete = TRUE,
.f = ~ any(.x, na.rm = TRUE)
) %>%
replace_na(FALSE) %>%
as.integer()
) %>%
ungroup()

saveRDS(data_nogeom, '/home/andreacatucci/output_new.RDS')
```

## Explore Data

### Distribution of solar generation (MW)
```{r eda1}
data_nogeom <- readRDS('/home/andreacatucci/output_new.RDS')

ggplot(data_nogeom, aes(x = generation_mw)) +
geom_histogram(bins = 100,
fill = "steelblue",
color = "white") +
# scale_x_log10() +  # optional: use if values are skewed
labs(title = "Distribution of Solar Generation (MW)")

cat(
"Number of observation with 0 generated energy: ",
data_nogeom %>% filter(generation_mw == 0) %>% count() %>% pull(n),
"\n"
)
```

#### Do we exclusively get 0 generation at night time?

```{r 0s}
hours_with_0s <- data_nogeom %>%
select(generation_mw, Hour, SUN_DUR) %>%
filter(generation_mw == 0) %>%
group_by(Hour) %>%
summarise(n = n(), SUN_DUR = mean(SUN_DUR))

hours_with_0s

gsps_with_0s <- data_nogeom %>%
select(generation_mw, Hour, gsp_id) %>%
filter(generation_mw == 0, !Hour %in% c(22, 23, 0, 01, 02, 03)) %>%
group_by(gsp_id) %>%
summarise(n = n()) %>%
arrange(desc(n))

gsps_with_0s

gsp_92_avg_gen <- data_nogeom %>%
filter(gsp_id == 92) %>%
select(generation_mw) %>%
summarise(mean(generation_mw))

gsp_92_avg_gen
```

#### We notice that there are 0 values for each hour actually. Night hours (10pm to 3am) have ONLY zero values, so we get rid of them and set those predictions equal to 0.
#### During the day, around 15 thousand observations per mid-day hour has 0 values, though the sun duration clearly is not 0 for most of them (look at average sunshine values). This means that there could have been some error in measurement or in the gsp itself.
#### However, we now notice that three gsps (92, 151 and 239) have only 0 values for generation, so we get rid of them from our data. GSP 149 only has data after 2022. ## Do we want to keep it for our modelling or not? This gsp corresponds to Bradford West, in the central part of the UK, slightly west of Leeds.


```{r found0}
data_nogeom <- data_nogeom %>%
filter(!gsp_id %in% c(92, 151, 239, 149), !Hour %in% c(22, 23, 0, 01, 02, 03))

ggplot(data_nogeom, aes(x = generation_mw)) +
geom_histogram(bins = 100,
fill = "steelblue",
color = "white") +
labs(title = "Distribution of Solar Generation (MW)")

cat(
"Number of observation with 0 generated energy: ",
data_nogeom %>% filter(generation_mw == 0) %>% count() %>% pull(n),
"\n"
)

hours_with_0s <- data_nogeom %>%
select(generation_mw, Hour, SUN_DUR) %>%
filter(generation_mw == 0) %>%
group_by(Hour) %>%
summarise(n = n(), SUN_DUR = mean(SUN_DUR))

hours_with_0s

gsps_with_0s <- data_nogeom %>%
select(generation_mw, Hour, gsp_id) %>%
filter(generation_mw == 0, !Hour %in% c(22, 23, 0, 01, 02, 03)) %>%
group_by(gsp_id) %>%
summarise(n = n()) %>%
arrange(desc(n))

gsps_with_0s

data_nogeom %>%
filter(gsp_id == 173) %>%
ggplot(aes(x = datetime_hour, y = generation_mw)) +
geom_line(color = "steelblue") +
labs(title = "Generation Over Time for GSP 173", x = "Datetime", y = "Generation (MW)") +
theme_minimal()
```

#### We reduced the number of observations during the day with 0 generation from 15000 to about 3500.

#### There still are around 1000 observations with 0 output during the day. We want to find out if these correspond to missing data or not, by grouping these observations.

```{r eda22222}
gsps_with_0s <- data_nogeom %>%
select(generation_mw, Hour, gsp_id, Year) %>%
filter(
generation_mw == 0,!Hour %in% c(16, 17, 18, 19, 20, 21, 22, 23, 0, 01, 02, 03, 04, 05, 06, 07, 08, 09, 10, 11)
) %>%
group_by(gsp_id, Year) %>%
summarise(n = n()) %>%
arrange(desc(n))

gsps_with_0s

data_nogeom %>%
filter(gsp_id == 173, Year %in% 2015:2018) %>%
ggplot(aes(x = datetime_hour, y = generation_mw)) +
geom_line(color = "steelblue") +
labs(title = "Generation Over 15-18 for GSP 173", x = "Datetime", y = "Generation (MW)") +
theme_minimal()

head(data_nogeom %>%
filter(
generation_mw == 0,!Hour %in% c(16, 17, 18, 19, 20, 21, 22, 23, 0, 01, 02, 03, 04, 05, 06, 07, 08, 09, 10, 11),
gsp_id == 173
))
```

#### 217 gsps have 0 values for generation during mid-day(12 to 3pm). Some gsps (173, 127, 316, 130, 11, 236,208, 337, 284, 333) have over 80 instances up to 765. It does not look though that there is systematic error.

#### When we look at gsp 173, we notice that mostly in winter months some values are 0 at times where you expect solar power generation. This gsp is located in Scotland (north of Edimburgh), so my assumption is that most of these values could be "real" and not missing data.

#### However, some values are strange. For example, "2015-06-27 15:00:00" has 0 generated energy, though weather variables seemed ideal for it. There are other values that are hard to explain through variables and could be measurement or plant errors. There seems to be no spatial explanation to these values.
```{r special case}
data_nogeom %>%
filter(datetime_hour %in% as.POSIXct(
c(
"2015-06-27 14:00:00",
"2015-06-27 15:00:00",
"2015-06-27 16:00:00"
)
), gsp_id == 173) %>%
select(
gsp_id,
generation_mw,
datetime_hour,
SUN_DUR,
CLD_TTL_AMT_ID,
MSL_PRESSURE,
AIR_TEMPERATURE,
PRCP_AMT,
MEAN_WIND_SPEED
)

gsps_with_zeros <- data_nogeom %>%
filter(
generation_mw == 0,!Hour %in% c(16, 17, 18, 19, 20, 21, 22, 23, 0, 01, 02, 03, 04, 05, 06, 07, 08, 09, 10, 11)
) %>%
group_by(gsp_id) %>%
summarise(total_zeros = n(), .groups = "drop")

gsp_locations <- data_nogeom %>%
select(gsp_id, Latitude, Longitude) %>%
distinct()

heatmap_data <- gsps_with_zeros %>%
inner_join(gsp_locations, by = "gsp_id")

ggplot(heatmap_data, aes(x = Longitude, y = Latitude)) +
geom_point(aes(size = total_zeros, color = total_zeros)) +
scale_color_viridis_c(option = "plasma") +
labs(
title = "Zero Generation Counts by GSP Location",
x = "Longitude",
y = "Latitude",
color = "Zero Counts",
size = "Zero Counts"
) +
theme_minimal()
```

### Replot distribution of solar power generation.

```{r eda123}
ggplot(data_nogeom, aes(x = generation_mw)) +
geom_histogram(bins = 100,
fill = "steelblue",
color = "white") +
labs(title = "Distribution of Solar Generation (MW)")

cat(
"Number of observation with 0 generated energy: ",
data_nogeom %>% filter(generation_mw == 0) %>% count() %>% pull(n),
"\n"
)
```

#### We have more than halved the 0 values. We probably should keep the remaining as we want the model to be able to predict 0 values (at 5am during different times of the year and with special weather conditions). We do still have a heavily skewed distribution, though.

### Numeric variables analysis

```{r eda2}
summary(select(data_nogeom, where(is.numeric)))
```
#### From this summary, we notice that the medians tend to be considered as not optimal weather for solar power generation (look sun duration, cloud amount).

#### We also notice that some variables are coherent with the skewness of data by looking at the mean-median relationship (SUN_DUR; CLD_AMT).

### Distributions of variables.

```{r eda4}
predictor_vars <- data_nogeom %>%
select(
SUN_DUR,
CLD_TTL_AMT_ID,
MSL_PRESSURE,
AIR_TEMPERATURE,
PRCP_AMT,
MEAN_WIND_DIR,
MEAN_WIND_SPEED,
MAX_GUST_DIR,
MAX_GUST_SPEED
) %>%
pivot_longer(cols = everything(),
names_to = "variable",
values_to = "value")

ggplot(predictor_vars, aes(x = value)) +
geom_histogram(
aes(y = ..density..),
bins = 30,
fill = "steelblue",
color = "white"
) +
facet_wrap(~ variable, scales = "free", ncol = 3) +
theme_minimal() +
labs(title = "Distribution of Numeric Predictors", x = "Value", y = "Density")
```

### Correlation plot

```{r eda3}
numeric_vars <- select(data_nogeom, where(is.numeric))
corr_matrix <- cor(numeric_vars, use = "complete.obs")
corrplot(
corr_matrix,
method = "color",
type = "lower",
tl.col = "black",
tl.cex = 0.8,
number.cex = 0.7,
diag = FALSE
)
```

#### This correlation plot shows the following:-generation of solar power is highly correlated with sun duration and correlated with air temperature.-Air temperature and sunshine duration are correlated between each other (likely that the former is influenced by the latter).
- Speed and direction of wind and gust are highly correlated between the two.
- Mean Sea Level Pressure is correlated with sun duration, air temperature, cloud coverage and wind speed (some existing literature says that higher MSL influences higher power generation).

### Time series plot
#### Plot of 15 randomly sampled gsp's complete time series.

```{r samples1}
set.seed(123)
sample_gsps <- sample(unique(data_nogeom$gsp_id), 15)

data_nogeom %>%
filter(gsp_id %in% sample_gsps) %>%
ggplot(aes(x = datetime_hour, y = generation_mw)) +
geom_line(color = "steelblue") +
labs(title = "Generation Over Time for 10 Random GSPs", x = "Datetime", y = "Generation (MW)") +
facet_wrap(~ gsp_id,
nrow = 5,
ncol = 3,
scales = "free") +
theme_minimal()
```

#### We notice strong seasonality throughout the year (summer-winter), and also an increasing trend (with different slopes for each gsp) throughout the years for most gsps. This is probably due to increased capacity (for which we do not have data prior to 2022). Could a proxy for capacity be the maximum value for each year for each gsp?

#### Day time series

```{r samples}
set.seed(12)
sample_gsps <- sample(unique(data_nogeom$gsp_id), 15)

data_nogeom %>%
filter(gsp_id %in% sample_gsps, Day == 31, Month == 5, Year == 2020) %>%
ggplot(aes(x = datetime_hour, y = generation_mw)) +
geom_line(color = "steelblue") +
labs(title = "Generation Over Time for 15 Random GSPs on 05/01/2020", x = "Datetime", y = "Generation (MW)") +
facet_wrap(~ gsp_id,
nrow = 5,
ncol = 3,
scales = "free") +
theme_minimal()

data_nogeom %>%
filter(gsp_id %in% sample_gsps, Day == 30, Month == 6, Year == 2017) %>%
ggplot(aes(x = datetime_hour, y = generation_mw)) +
geom_line(color = "steelblue") +
labs(title = "Generation Over Time for 15 Random GSPs on 06/30/2017", x = "Datetime", y = "Generation (MW)") +
facet_wrap(~ gsp_id,
nrow = 5,
ncol = 3,
scales = "free") +
theme_minimal()

data_nogeom %>%
filter(gsp_id %in% sample_gsps, Day == 10, Month == 9, Year == 2015) %>%
ggplot(aes(x = datetime_hour, y = generation_mw)) +
geom_line(color = "steelblue") +
labs(title = "Generation Over Time for 15 Random GSPs on 09/10/2015", x = "Datetime", y = "Generation (MW)") +
facet_wrap(~ gsp_id,
nrow = 5,
ncol = 3,
scales = "free") +
theme_minimal()
```

#### On a daily basis, the trend is of course bell-shaped but it can show lots of differences in the spikes and slopes, probably depending on weather conditions.

#### Further analysis can be done on the 0 values. We first check if we are able to cluster these values by gsp and season. Also, we try modelling logistic regression on the anomaly binary variable to observe unexpected anomalies.

```{r eda6}
data_nogeom <- data_nogeom %>%
mutate(
anomaly_zero_generation = if_else(generation_mw == 0 &
SUN_DUR > 0.2, 1, 0),
Season = case_when(
Month %in% c(12, 1, 2) ~ "Winter",
Month %in% c(3, 4, 5) ~ "Spring",
Month %in% c(6, 7, 8) ~ "Summer",
TRUE ~ "Autumn"
)
)

anomaly_counts <- data_nogeom %>%
group_by(gsp_id, Season) %>%
summarise(total_anomalies = sum(anomaly_zero_generation),
.groups = "drop")

gsp_locations <- data_nogeom %>%
select(gsp_id, Latitude, Longitude) %>%
distinct()

plot_data <- inner_join(anomaly_counts, gsp_locations, by = "gsp_id")

ggplot(plot_data, aes(x = Longitude, y = Latitude)) +
geom_point(aes(size = total_anomalies, color = total_anomalies)) +
facet_wrap(~ Season) +
scale_color_viridis_c(option = "plasma") +
labs(
title = "Anomalous Zero Generation Counts by GSP and Season",
x = "Longitude",
y = "Latitude",
color = "Anomaly Count",
size = "Anomaly Count"
) +
theme_minimal()
```

#### It does not seem like we can cluster by season and location.

## Further ideas:
- Remove all night 0s by looking at sunrise and sunset times for each month.
- How to model information about wind and clouds to understand cloud movements correlated to generation.
- Flag anomalies with a binary variable.
- Create interaction variables (e.g. good_weather = [sunshine > 0.5, clouds < 0.2])
- Use each year's maximum value as capacity for that year.

## Remove night 0s by obtaining sunrise and sunset times.

#### We use the suncalc package to obtain sunrise and sunset times for each gsp location for each day, and apply this to every year in data to only keep meaningful values.

```{r suncalc, include = F}
gsp_locations <- data_nogeom %>%
select(gsp_id, lon=Longitude, lat=Latitude)

dates <- seq(as.Date("2020-01-01"), as.Date("2020-12-31"), by = "day")

gsp_dates <- crossing(gsp_locations, date = dates)

sun_times <- getSunlightTimes(
data = gsp_dates,
keep = c("sunrise", "sunset"),
tz = "Europe/London"
)

sun_times$gsp_id <- gsp_dates$gsp_id

sun_bounds <- sun_times %>%
mutate(
Month = month(sunrise),
Day = day(sunrise),
sunrise_hour = hour(ceiling_date(sunrise, unit = "hour")),
sunset_hour  = hour(ceiling_date(sunset, unit = "hour"))
) %>%
select(gsp_id, Month, Day, sunrise_hour, sunset_hour)

filtered_data <- data_nogeom %>%
left_join(sun_bounds, by = c("gsp_id", "Month", "Day")) %>%
filter(Hour >= sunrise_hour, Hour <= sunset_hour)

saveRDS(filtered_data, '/home/andreacatucci/filtered_data.RDS')
```

### Average seasonal plots

```{r avgseas}
filtered_data %>%
mutate(
Season = case_when(
Month %in% c(12, 1, 2) ~ "Winter",
Month %in% c(3, 4, 5) ~ "Spring",
Month %in% c(6, 7, 8) ~ "Summer",
TRUE ~ "Fall"
)
) %>%
group_by(Hour, Season) %>%
summarise(mean_gen = mean(generation_mw, na.rm = TRUE)) %>%
ggplot(aes(x = Hour, y = mean_gen, color = Season)) +
geom_line() +
labs(title = "Average Diurnal Generation Curve by Season")
```

### Check spatial correlation with anomalies.

```{r anomaly1}
gsps_with_zeros <- filtered_data %>%
filter(generation_mw == 0) %>%
group_by(gsp_id) %>%
summarise(total_zeros = n(), .groups = "drop") %>%
arrange(desc(total_zeros))

gsps_with_zeros

gsp_locations <- filtered_data %>%
select(gsp_id, Latitude, Longitude) %>%
distinct()

heatmap_data <- gsps_with_zeros %>%
inner_join(gsp_locations, by = "gsp_id")

ggplot(heatmap_data, aes(x = Longitude, y = Latitude)) +
geom_point(aes(size = total_zeros, color = total_zeros)) +
scale_color_viridis_c(option = "plasma") +
labs(
title = "Zero Generation Counts by GSP Location",
x = "Longitude", y = "Latitude",
color = "Zero Counts", size = "Zero Counts"
) +
theme_minimal()
```
#### It looks like gsps in the South count more anomalies (with gsp 173 in the North being an outlier).

### Fit lightgbm model to find anomalies.

```{r anomaly2}
filtered_data <- filtered_data %>%
mutate(generation_zero = if_else(generation_mw == 0, 1, 0))

features <- c("SUN_DUR", "CLD_TTL_AMT_ID", "MSL_PRESSURE", "AIR_TEMPERATURE",
"PRCP_AMT", "MEAN_WIND_SPEED", "MAX_GUST_SPEED", "HAS_RAINED", "Latitude", "Longitude", "Elevation", "Hour",
"Year", "Month", "Quarter")

X <- data.matrix(filtered_data[, features])
y <- filtered_data$generation_zero

dtrain <- lgb.Dataset(X, label = y)

params <- list(
objective = "binary",
metric = "binary_logloss",
num_leaves = 31,
learning_rate = 0.1,
verbosity = -1
)

lgb_model <- lgb.train(params, dtrain, nrounds = 100)

filtered_data$prob_zero <- predict(lgb_model, X)

filtered_data <- filtered_data %>%
mutate(unexpected_zero = if_else(generation_mw == 0 & prob_zero < 0.5, 1L, 0L))

unexpected_summary <- filtered_data %>%
filter(unexpected_zero == 1) %>%
group_by(gsp_id, Latitude, Longitude) %>%
summarise(unexpected_count = n(), .groups = "drop") %>%
arrange(desc(unexpected_count))

unexpected_summary

ggplot(unexpected_summary, aes(x = Longitude, y = Latitude)) +
geom_point(aes(size = unexpected_count, color = unexpected_count)) +
scale_color_viridis_c(option = "inferno") +
labs(title = "Unexpected Zero Generation by GSP",
x = "Longitude", y = "Latitude") +
theme_minimal()

filtered_data %>%
filter(unexpected_zero == 1)
```

#### GSP 173 counts many more unexpected 0s compared to other gsps.
 
```{r gsp173}
filtered_data %>%
filter(unexpected_zero == 1, gsp_id==173)
```


#### Notice how most of "unexpected zeros" could actually be expected given time of the day is right after sunrise or right before sunset and weather conditions are bad (0 sunshine).

```{r groupseason2}
filtered_data %>%
filter(unexpected_zero == 1) %>%
count(Hour) %>%
ggplot(aes(x=Hour, y=n, fill=Hour)) +
geom_col() +
labs(title = "Total Unexpected Zeros by Hour",
x = "Hour", y = "Count") +
theme_minimal()

filtered_data %>%
filter(unexpected_zero == 1) %>%
count(Season) %>%
ggplot(aes(x = Season, y = n, fill = Season)) +
geom_col() +
labs(title = "Total Unexpected Zeros by Season",
x = "Season", y = "Count") +
theme_minimal()

top_gsps <- filtered_data %>%
filter(unexpected_zero == 1) %>%
count(gsp_id, name = "total_unexpected") %>%
top_n(20, total_unexpected) %>%
pull(gsp_id)

filtered_data %>%
filter(unexpected_zero == 1, gsp_id %in% top_gsps) %>%
group_by(gsp_id, Season) %>%
summarise(unexpected_count = n(), .groups = "drop") %>%
ggplot(aes(x = Season, y = unexpected_count, fill = Season)) +
geom_bar(stat = "identity", position = "dodge") +
facet_wrap(~ gsp_id, scales = "free_y") +
labs(title = "Top 20 GSPs: Unexpected Zeros by Season") +
theme_minimal()
```

#### We notice that the gsp with the most 0 values are also those with the most unexpected 0 values. This means that those 0 values are probably due to measurements erros. Again, no spatial explanation of missing values. Do we remove these values from our dataset?
#### The last plots also show that the classifier has a hard time classifying as 0 timestamps right at sunrise or sunset.

#### What was correctly classified as 0?


```{r real}
filtered_data <- filtered_data %>%
mutate(expected_zero = if_else(generation_mw == 0 & prob_zero > 0.5, 1L, 0L))

expected_summary <- filtered_data %>%
filter(expected_zero == 1) %>%
group_by(gsp_id, Latitude, Longitude) %>%
summarise(expected_count = n(), .groups = "drop") %>%
arrange(desc(expected_count))

filtered_data %>%
filter(expected_zero == 1)
```

#### This model has a recall of .72 on the training data meaning that 28% are "anomalies". As we can see from the data, these anomalies are not clearly anomalies, and do not differ much from the correct classifications.

```{r featimp}
lgb.importance(lgb_model) %>%
as.data.frame() %>%
ggplot(aes(x = reorder(Feature, Gain), y = Gain)) +
geom_col(fill = "steelblue") +
coord_flip() +
labs(title = "Feature Importance for Predicting Zero Generation")

set.seed(123)

features_scaled <- unique(scale(filtered_data[, features]))
pca <- prcomp(features_scaled, center = TRUE, scale. = TRUE)
summary(pca)  
loadings <- pca$rotation
head(loadings[, 1:8]) 

filtered_data <- filtered_data %>%
mutate(
near_sunrise_sunset = Hour < (sunrise_hour + 1) | Hour > (sunset_hour - 1),
low_sunshine = SUN_DUR < 0.1
)

filtered_data %>%
filter(unexpected_zero == 1) %>%
count(near_sunrise_sunset, low_sunshine)
```

#### ~84% (113,524 / 184,613) of unexpected zeros occurred when both conditions are met: near sunrise/sunset and low sunshine. These are very likely non-anomalous edge cases. Something we want the model to predict to be 0.
#### Only ~8,000 (~4%) seem genuinely anomalous (FALSE, FALSE). This amounts to 0.0005% of the data.

```{r more_anomalies}
filtered_data %>%
  filter(near_sunrise_sunset==F,low_sunshine==F,generation_mw==0) %>%
  group_by(gsp_id) %>%
  summarise (n = n()) %>%
  arrange(desc(n))

filtered_data %>%
  filter(near_sunrise_sunset==F,low_sunshine==F,generation_mw==0, gsp_id==173)

filtered_data <- filtered_data %>%
  mutate(Anomaly = if_else(
    near_sunrise_sunset == FALSE & low_sunshine == FALSE & generation_mw == 0,
    1L,  
    0L   
  ))

saveRDS(filtered_data, '/home/andreacatucci/filtered_data.RDS')
```
#### This is what the anomalies look like.

#### How to deal with them? First, flag them with a binary variable.
#### Depending on model, we can choose different approaches. For initial lineare regression, we can just get rid of these values as we do not need continuity. In temporal and spatial approaches, we must keep them in some way (interpolation?). In deep learning, we can use the flag.

### Distribution of solar generation (MW) in filtered dataset
```{r eda1}
ggplot(filtered_data, aes(x = generation_mw)) +
geom_histogram(bins = 100,
fill = "steelblue",
color = "white") +
labs(title = "Distribution of Solar Generation (MW)")

cat(
"Number of observation with 0 generated energy: ",
filtered_data %>% filter(generation_mw == 0) %>% count() %>% pull(n),
"\n"
)
```


### Linear Regression

#### We will try fitting linear regression on the dataset on Python.
#### We prepare our data.

```{r lin}
lin_regr_data <- filtered_data %>%
  filter(Anomaly==0) %>%
  select(-c(gsp_id, anomaly_zero_generation, Quarter, generation_zero, prob_zero, unexpected_zero, expected_zero, Anomaly))

write_feather(lin_regr_data, '/home/andreacatucci/lin_regr_data.feather')
```

### Distribution of solar generation (MW) in filtered dataset
```{r eda1}
ggplot(lin_regr_data, aes(x = scale(generation_mw))) +
geom_histogram(bins = 100,
fill = "steelblue",
color = "white") +
labs(title = "Distribution of Solar Generation (MW)")
```

#### First linear regression model obtains an R2 of 0.085 and an MSE of 1.24 on standardized data. This model uses as predictors all those in the dataset with no interaction terms. 
#### Some issues with the model that lead to poor results:
- Solar power generation can be highly non-linear with respect to weather variables (sun duration = 0 means 0 generation, hour and month has non linear effect when turned into a sin cosine function, etc.).
- No interaction terms were added.
- Distribution of generation is far from normal.

#### We try adding interaction terms.

- SUN_DUR_x_Month
- SUN_DUR_x_hour_cos
- SUN_DUR_x_CLD_TTL_AMT_ID
- SUN_DUR_x_Longitude
- SUN_DUR_x_hour_cos_x_month_cos

#### We obtain an R2 of 0.089, 4% better then the previous model. However, performance is still poor probably due to other issues stated.